\chapter{Project Proposal \\
\small{\textit{-- Justin Baumann, Gianna Cerbone, Thomas Ung, Spurthi Setty}} 
\index{Chapter!ProjectProposal}
\index{ProjectProposal}
\label{Chapter::ProjectProposal}}

% Add a section and label it so that we can reference it later
\section{Project Proposal \label{Section:Description}}

\subsection{Project Title: QuackOps User Interface}

The QuackOps Senior Design project has the following mission statement: 
To develop an autonomous drone delivery system that uses AI for navigation and visual target recognition, and to provide fast, contactless, and efficient on-campus delivery of goods and food.

To aid in the QuackOps Senior Design project, our team will help in developing the graphic user interface with computer vision, AI, and real time updating components from the data gathered by the QuackOps drone. Our contribution will be the web based dashboard that allows users to define delivery parameters, track the drone's location in real time, manage geofences, and monitor fleet health and compliance logs. Our main focus will be the software and user centered interface using DevOps techniques and skills learned in class. 

We will be using Jira KANBAN for task tracking and distribution. Github for source control. CI/CD Actions within GitHub is what we are considering for testing as it is implemented directly into GitHub. Our team needs to make sure that the interface is modular and can be integrated with whatever we are using to get the drone data and can also be able to communicate with the drone in some way. For this part - we will need to think about the drone project extensively and keep both projects intertwined. 


